---
title: "covid-p"
author: "Sara"
date: "11/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(forecast)
library(tidyverse)
library(kableExtra)
library(here)
library(plyr)
library(dplyr)
library(fastDummies)
library(caTools)
library(gains)
library(caret)
library(MASS)
library(FNN)
```



```{r}
t.covid.df <- read.csv("total.csv", header = TRUE)
head(t.covid.df)
```

# loading libararies and dataset

```{r}
dim(t.covid.df)
```

```{r}
str(t.covid.df )
```

```{r}
summary(t.covid.df)
```

```{r}
newdata <- t.covid.df
newdata.df <- mutate_if(newdata, is.character, as.factor)
str(newdata.df)
```


```{r}
summary(newdata.df)
```
# looking into variables and its structure
```{r}
table(newdata.df$death_yn, newdata.df$medcond_yn)
```

```{r}
table(newdata.df$death_yn, newdata.df$icu_yn)
```

```{r}
table(newdata.df$death_yn, newdata.df$hosp_yn)
```

#lokking into regressors by outcome variable and revaluing catagories of regressors 

```{r}

newdata.df$hosp_yn <- revalue (newdata.df$hosp_yn, c ( "Missing" = "No", "Unknown" = "No"))
summary(newdata.df$hosp_yn)
```

```{r}
table(newdata.df$death_yn, newdata.df$hosp_yn)
```


```{r}
table(newdata.df$death_yn, newdata.df$current_status)

```

```{r}
table(newdata.df$death_yn, newdata.df$age_group)
ggplot(data = newdata.df) +
  geom_bar(mapping = aes(x = age_group))
```

```{r}
newdata.df$Age <- revalue(newdata.df$age_group, c("0 - 9 Years"="19 & Below","10 - 19 Years"= "19 & Below", "20 - 29 Years"= "20 - 39 Years", "30 - 39 Years"= "20 - 39 Years", "Unknown"= "20 - 39 Years" ,"40 - 49 Years"= "40 - 59 Years", "50 - 59 Years"= "40 - 59 Years", "60 - 69 Years"= "60 -79","70 - 79 Years"= "60 -79", "80+ Years"= "80+ Years" ))
```

```{r}
table(newdata.df$death_yn, newdata.df$Age)
```

```{r}
table(newdata.df$death_yn, newdata.df$Race.and.ethnicity..combined.)
```



```{r}
newdata.df$Ethnicity <- revalue(newdata.df$Race.and.ethnicity..combined., c("Hispanic/Latino" = "Hispanic" , "American Indian/Alaska Native, Non-Hispanic" = "Non-Hispanic", "Black, Non-Hispanic" = "Non-Hispanic", "Multiple/Other, Non-Hispanic" = "Non-Hispanic", "Asian, Non-Hispanic" = "Non-Hispanic", "Native Hawaiian/Other Pacific Islander, Non-Hispanic" = "Non-Hispanic", "White, Non-Hispanic" = "Non-Hispanic" , "Unknown" = "Non-Hispanic" ))
```

```{r}
table(newdata.df$death_yn, newdata.df$Ethnicity)
```

```{r}
newdata.df$Race <- revalue(newdata$Race.and.ethnicity..combined., c("Hispanic/Latino" = "Others" , "American Indian/Alaska Native, Non-Hispanic" = "Asian", "Black, Non-Hispanic" = "Black", "Multiple/Other, Non-Hispanic" = "Others", "Asian, Non-Hispanic" = "Asian", "Native Hawaiian/Other Pacific Islander, Non-Hispanic" = "Others", "White, Non-Hispanic" = "White", "Unknown" = "Others"))
```

```{r}
table(newdata.df$death_yn, newdata.df$Race)

```

```{r}
newdata.df$Race2 <- revalue(newdata.df$Race, c("Asian" = "White/Black/Asian" , "Black" = "White/Black/Asian", "White" = "White/Black/Asian", "Others"= "Others"))


newdata.df$Race2  <- ifelse(newdata.df$Race  == "Others", 1, 0)
table(newdata.df$death_yn, newdata.df$Race2)

```


```{r}
table(newdata.df$death_yn, newdata.df$sex)
```

```{r}
newdata.df$Gender<- revalue(newdata$sex, c("Missing"= "Female", "Unknown" = "Female"))
```

```{r}
table(newdata.df$death_yn, newdata.df$Gender)
```


```{r}
newdata.df$Month<-months(as.Date(newdata.df$cdc_report_dt))
newdata.df$Weekday<-weekdays(as.Date(newdata.df$cdc_report_dt))
```

```{r}
table(newdata.df$death_yn, newdata.df$Month)
```

```{r}
table(newdata.df$Month)
```


```{r}
ggplot(data = newdata.df) +
  geom_bar(mapping = aes(x = Age),  fill="pink", color="purple")
ggplot(data = newdata.df) +
  geom_bar(mapping = aes(x = Gender), fill="Blue", color="purple" )
ggplot(data = newdata.df) +
  geom_bar(mapping = aes(x = Race), fill="pink", color="pink" )
ggplot(data = newdata.df) +
  geom_bar(mapping = aes(x = Ethnicity), fill="red", color="purple")
ggplot(data = newdata.df) +
  geom_bar(mapping = aes(x = Month),  fill="purple", color="purple")
```

```{r}
newdata.df$Months<- revalue(newdata.df$Month, c("February"= "February-March", "March" = "February-March"))
table(newdata.df$death_yn, newdata.df$Months)
```


```{r}
newdata.df$onset_dt <- ifelse(newdata.df$onset_dt == "", 0, 1)
table(newdata.df$death_yn, newdata.df$onset_dt)
```



```{r}
summary(newdata.df)
```

```{r}
table(newdata.df$Ethnicity)
```

```{r}
levels(newdata.df$Ethnicity)
```
```{r}
ggplot(newdata.df, aes(y = death_yn, x = current_status, colour= Ethnicity)) +
geom_point(alpha = 0.6)
ggplot(newdata.df, aes(y = death_yn, x = current_status, colour= Gender)) +
geom_point(alpha = 0.6)

```


```{r}
newdata2.df <- newdata.df
newdata2.df$Gender <- ifelse(newdata2.df$Gender == "Female", 1, 0)
newdata2.df$hosp_yn <- ifelse(newdata2.df$hosp_yn == "No", 1, 0)
newdata2.df$current_status<- ifelse(newdata2.df$current_status == "Laboratory-confirmed case", 1, 0)
newdata2.df$death_yn<- ifelse(newdata2.df$death_yn == "Yes", 1, 0)
newdata2.df$Ethnicity<- ifelse(newdata2.df$Ethnicity == "Non-Hispanic", 1, 0)
head(newdata2.df)
```



```{r}
newdata2.df$AGE1 <- as.factor(revalue(newdata2.df$Age, c("19 & Below"= "0","20 - 39 Years"="0", "40 - 59 Years"= "0", "60 -79"="1",  "80+ Years" ="1")))
levels(newdata2.df$AGE1)
```



```{r}

table(newdata2.df$AGE1)

```


```{r}
table(newdata2.df$Ethnicity)
```

```{r}
names(newdata2.df)
```

```{r}
data_d.df <- newdata2.df[c(1,4,5,9,11,13:17,20:21)]
str(data_d.df)
```

```{r}
data_d.df<- mutate_if(data_d.df, is.character, as.factor)
```

```{r}
library(fastDummies)
results <- fastDummies::dummy_cols(data_d.df, select_columns = c( 'Age', 'Race'))
```

```{r}
head(results)
```
```{r}
t(t(names(results)))
```

#Transformed variables and created dummies 

```{r}
finaldata.df<- results [c(1:5,7,9:12,14:17, 19:21)]
dim(finaldata.df)
names(finaldata.df)
```

#Created newdataset of transformed variables

```{r}
finaldata.df<- mutate_if(finaldata.df, is.numeric, as.integer)
finaldata.df<- mutate_if(finaldata.df, is.factor, as.integer)
str(finaldata.df)
```
```{r}
colnames(finaldata.df)[11] <- c("Age_20to39Years")
colnames(finaldata.df)[12] <- c("Age_40to59Years")
colnames(finaldata.df)[13] <- c("Age_60to79Years")
colnames(finaldata.df)[14] <- c("Age_80plus")
```


```{r}
str(finaldata.df)
```

```{r}
set.seed(1)
train.rows <- sample(rownames(finaldata.df), dim(finaldata.df)[1]*0.6)
train.data <- finaldata.df[train.rows, ]
```

```{r}
str(train.data)
```


```{r}
dim(train.data)
```

```{r}
valid.rows <- setdiff(rownames(finaldata.df), train.rows)
valid.data <- finaldata.df[valid.rows, ]
head(valid.data)
dim(valid.data)
```
```{r}
t(t(names(train.data)))
```
 # partitioned Dataset into test and train 

```{r}
logitMod <- glm(death_yn ~ onset_dt+current_status+Gender+hosp_yn+Ethnicity+Age_20to39Years+Age_40to59Years+Age_60to79Years+Age_80plus+Race_Black+Race_Others+Race_White, data=train.data, family=binomial)
options(scipen = 999)
summary(logitMod)
```

```{r}
logitMod.pred <- predict.glm(logitMod, valid.data , type = "response")
opslm1 <- as.factor(ifelse(logitMod.pred>0.5, "1", "0"))

confusionMatrix(as.factor(valid.data$death_yn), opslm1, positive = "1")
```

# Run logistc Model
```{r}
table(newdata2.df$Age, newdata$death_yn)
```



```{r}
stepwise_logit_mod <- glm(death_yn ~ onset_dt+Gender+current_status+hosp_yn+Ethnicity+Race2+Race_Black+Race_Others+Race_White+ AGE1, data=train.data, family=binomial) %>% stepAIC(trace=TRUE)
```


```{r}
summary(stepwise_logit_mod)
```

```{r}
stepwise_logit_mod.pred <- predict.glm(stepwise_logit_mod, valid.data , type = "response")
opslm <- as.factor(ifelse(stepwise_logit_mod.pred>0.5, "1", "0"))

confusionMatrix(as.factor(valid.data$death_yn), opslm, positive = "1")
hist(stepwise_logit_mod.pred)
```

# run Stepwise logistics model to finalize the model with right number of signicant regressors

```{r}
ba <- c()
sen <- c()
spe <- c()
co <- c()
for (i in c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9)){
  op_i <- as.factor(ifelse(stepwise_logit_mod.pred>i, "1", "0"))
  cm <- confusionMatrix(as.factor(valid.data$death_yn), op_i, positive = "1")
  co <- c(co, i)
  ba <- c(ba, cm$byClass[11])
  sen <- c(sen, cm$byClass[1])
  spe <- c(spe, cm$byClass[2])
}
```




```{r}
result.slm.df <- data.frame(co, ba, sen, spe)
result.slm.df
```

# created table ofscores of  validity tests of model
```{r}
write.csv(result.slm.df , "results.slm.csv")
```



```{r}
opslm.h <- as.factor(ifelse(stepwise_logit_mod.pred>0.7, "1", "0"))
confusionMatrix(as.factor(valid.data$death_yn), opslm.h, positive = "1")
```

```{r}
opslm.h <- as.factor(ifelse(stepwise_logit_mod.pred>0.7, "1", "0"))
confusionMatrix(as.factor(valid.data$death_yn), opslm.h, positive = "1")
```


```{r}
data.frame(coeff = stepwise_logit_mod$coefficients, odds = exp(stepwise_logit_mod$coefficients))
```

```{r}
paste("P(Not Survived)= 1/1+e^-(-2.272 -1.323(onset_dt)-0.397(Gender)-1.207(current_status)-2.453(hosp_yn) -0.732(Ethnicity)-0.759(Race)+ 2.772 (AGE1)")
```


```{r}
logitMod.3 <- glm(death_yn ~ onset_dt+current_status+hosp_yn+Ethnicity+Race_Black+Race_Others+Race_White+ AGE1, data=train.data, family=binomial)
options(scipen = 999)
summary(logitMod.3)
```



```{r}
logitMod.3.pred <- predict.glm(logitMod.3, valid.data , type = "response")
op3 <- as.factor(ifelse(logitMod.3.pred>0.5, "1", "0"))

confusionMatrix(as.factor(valid.data$death_yn), op3, positive = "1")
hist(logitMod.3.pred)
```

```{r}
op3.l <- as.factor(ifelse(logitMod.3.pred>0.3, "1", "0"))
confusionMatrix(as.factor(valid.data$death_yn), op3.l, positive = "1")
```

```{r}
op3.h <- as.factor(ifelse(logitMod.3.pred>0.7, "1", "0"))
confusionMatrix(as.factor(valid.data$death_yn), op3.h, positive = "1")
```



```{r}
data.frame(coeff = logitMod.3$coefficients, odds = exp(logitMod.3$coefficients))
```



```{r}
summary(logitMod.3, odds = exp(logitMod.3$coefficients))
```

```{r}
ba <- c()
sen <- c()
spe <- c()
co <- c()
for (i in c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9)){
  op_i <- as.factor(ifelse(logitMod.3.pred>i, "1", "0"))
  cm <- confusionMatrix(as.factor(valid.data$death_yn), op_i, positive = "1")
  co <- c(co, i)
  ba <- c(ba, cm$byClass[11])
  sen <- c(sen, cm$byClass[1])
  spe <- c(spe, cm$byClass[2])
}
```

```{r}
result_df <- data.frame(co, ba, sen, spe)
```

```{r}
result_df
```


```{r}
write.csv(result_df, "results.csv")
```

```{r}
plot(result_df$ba, result_df$co, bty="l", lty=1)
```

```{r}
gain <- gains(valid.data$death_yn, stepwise_logit_mod.pred, groups=length(stepwise_logit_mod.pred))
# plot lift chart
plot(c(0,gain$cume.pct.of.total*sum(valid.data$death_yn))~c(0,gain$cume.obs),
xlab="# cases", ylab="Cumulative", main="", type="l")
lines(c(0,sum(valid.data$death_yn))~c(0, dim(valid.data)[1]), lty=2)
# compute deciles and plot decile-wise chart
gain <- gains(valid.data$death_yn, stepwise_logit_mod.pred, groups=10)
heights <- gain$mean.resp/mean(valid.data$death_yn)
midpoints <- barplot(heights, names.arg = gain$depth, ylim = c(0,9),
xlab = "Percentile", ylab = "Mean Response", main = "Decile-wise lift chart")
# add labels to columns
text(midpoints, heights+0.5, labels=round(heights, 1), cex = 0.8)
```


```{r}
## A 3-nearest neighbours model with no normalization
ba_nn <- c()
sen_nn <- c()
spe_nn <- c()
k <- c()
for (i in c(1, 2, 3, 4, 5, 6, 7, 8, 9)){
  nn <- DMwR::kNN(death_yn ~ onset_dt+current_status+Gender+hosp_yn+Ethnicity+Race_Black+Race_Others+Race_White+ AGE2, train = train.data, test = valid.data, norm=TRUE,k=i)
  cm <- confusionMatrix(as.factor(valid.data$death_yn), nn, positive = "1")
  k <- c(k, i)
  ba_nn <- c(ba_nn, cm$byClass[11])
  sen_nn <- c(sen_nn, cm$byClass[1])
  spe_nn <- c(spe_nn, cm$byClass[2])
}
```

#validity test scores for multiple logistic models

```{r}
head(train.data, 1)
```

```{r}
# 2, 3, 4, 5, 6, 7, 10,  16, 
temp.train <- subset(train.data, select = c(2, 3, 4,5, 6, 7,9))
temp.valid <- subset(valid.data, select = c(2, 3, 4, 5,6, 7, 9))
```



```{r}
nn <- FNN::knn(train = temp.train, test = head(temp.valid, 1), cl = temp.train$death_yn, k = 3)
```

```{r}
nn
```

```{r}
head(temp.valid, 1)
```

```{r}
as.integer(row.names(temp.train)[attr(nn, "nn.index")])
```

```{r}
temp.train[c("466", "1104", "471"), ]
```

# used K-Nearest neighbour test where K=3 



```{r}
results_nn_df <- data.frame(k, ba_nn, sen_nn, spe_nn)
results_nn_df
```

#validation score of K_NN model

```{r}
write.csv(results_nn_df, "results_nn.csv")
```




